\Large Exponential Random Graph Models with Big Networks: Maximum Pseudolikelihood Estimation and the\\Parametric Bootstrap

### 2017-05-16 TeX/paper.tex

We estimate the coefficient of this matrix using both techniques, the MCMLE and the bootstrapped MPLE. The MCMLE requires a sample size of at least $100000$ networks to converge. The bootstrapped MPLE was estimated by using 500 simulated networks. As we described in the chapter {\it Estimation}, only one edge at a time is changed when simulating networks. For better comparison, we chose the same Burn-In ($163840$ MH-steps) and the same number of iterations ($10240$ MH-steps) for sampling networks. The results can be found in table ??.
$$\text{insert table with boot MPLE and MCMLE results}$$
The bootstrapped MPLE is not only simple and fast, it is also embarrassingly parallel, meaning that there is no problem seperating the computation into several tasks. In other words, by using multiple cores, the computing time for estimating bootstrapped MPLE confidence intervals can be reduced substantially. Figure \ref{comptime} illustrates the relative computing time of the bootstrapped MPLE using 500 simulated networks and the MCMLE for the three networks Faux Mesa High (205 nodes), Faux Magnolia High (1461 nodes) and Supreme Court Citation (9223 nodes) for an increasing number of computing cores. For the small network we simulate $2000$ networks using a MCMC interval of $2000$ steps, for the medium network we simulate $8000$ networks using a MCMC interval of $5000$ steps and for the large network we simulate 1 million networks using $10,000$ MCMC steps in order to approximate the likelihood appropriately. The chosen sample sizes and MCMC steps are necessary to guarantee a good model fit. We define the simulation time of the bootstrapped MPLE as a function of the number of available computing cores x:

>>>>

We estimate the coefficient of this matrix using both techniques, the MCMLE and the bootstrapped MPLE. The MCMLE requires a sample size of at least $100000$ networks to converge. The bootstrapped MPLE was estimated by using 500 simulated networks. As we described in the chapter {\it Estimation}, only one edge at a time is changed when simulating networks. For better comparison, we chose the same Burn-In ($163840$ MH-steps) and the same number of iterations ($10240$ MH-steps) for sampling networks. The results can be found in table \ref{sc_results}.\\[0.3cm]
\begin{table}[]
\begin{tabular}{|
>{\columncolor[HTML]{EFEFEF}}l |l|l|l|l|l|l|}
\hline
\cellcolor[HTML]{EFEFEF}                   & \multicolumn{2}{c|}{\cellcolor[HTML]{EFEFEF}MCMLE}                   & \multicolumn{2}{c|}{\cellcolor[HTML]{EFEFEF}bootstrapped MPLE}       & \multicolumn{2}{c|}{\cellcolor[HTML]{EFEFEF}Logistic Regression}     \\ \cline{2-7} 
\multirow{-2}{*}{\cellcolor[HTML]{EFEFEF}} & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}St. Error & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}St. Error & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}St. Error \\ \hline
Edges                                      & -5.657                           & 0.028                             &                                  &                                   & -4.568                           & 0.038                             \\ \hline
Nodecov.Salience                           & 0.592                            & 0.012                             &                                  &                                   & 1.482                            & 0.009                             \\ \hline
Nodecov.Year                               & 0.253                            & 0.012                             &                                  &                                   & 0.208                            & 0.015                             \\ \hline
Absdiff.Year                               & -2.470                           & 0.160                             &                                  &                                   & -3.121                           & 0.194                             \\ \hline
Absdiff.Year.Squared                       & 0.221                            & 0.066                             &                                  &                                   & 0.436                            & 0.080                             \\ \hline
DSP(0)                                     & 0.0005                           & 0.0002                            &                                  &                                   & 0.021                            & 0.0002                            \\ \hline
ESP(0)                                     & -3.129                           & 0.012                             &                                  &                                   & -2.162                           & 0.004                             \\ \hline
\end{tabular}
\caption{Estimation results for the Supreme Court Network using MCMLE, bootstrapped MPLE and logistic regression}
\label{sc_results}
\end{table}
\noindent The bootstrapped MPLE is not only simple and fast, it is also embarrassingly parallel, meaning that there is no problem seperating the computation into several tasks. In other words, by using multiple cores, the computing time for estimating bootstrapped MPLE confidence intervals can be reduced substantially. Figure \ref{comptime} illustrates the relative computing time of the bootstrapped MPLE using 500 simulated networks and the MCMLE for the three networks Faux Mesa High (205 nodes), Faux Magnolia High (1461 nodes) and Supreme Court Citation (9223 nodes) for an increasing number of computing cores. For the small network we simulate $2000$ networks using a MCMC interval of $2000$ steps, for the medium network we simulate $8000$ networks using a MCMC interval of $5000$ steps and for the large network we simulate 1 million networks using $10,000$ MCMC steps in order to approximate the likelihood appropriately. The chosen sample sizes and MCMC steps are necessary to guarantee a good model fit. We define the simulation time of the bootstrapped MPLE as a function of the number of available computing cores x:
